---
title: "Report"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)

src_url = "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
data = read.csv(src_url)
```

## Intentions
The purpose of the analysis is to get a general idea of the timing of these incidents and where they took place most frequently.


## Data transformations and initial summary

```{r transform, echo=TRUE}
data = select(data, c("OCCUR_DATE", "OCCUR_TIME", "BORO", "LOCATION_DESC"))

data$OCCUR_DATETIME = str_c(data$OCCUR_DATE, " ", data$OCCUR_TIME)
data = select(data, -c("OCCUR_DATE", "OCCUR_TIME"))
data$OCCUR_DATETIME = strptime(data$OCCUR_DATETIME, "%m/%d/%Y %H:%M:%S")
#data$OCCUR_DATETIME$zone = NULL

data$BORO = as.factor(data$BORO)
data$LOCATION_DESC = as.factor(data$LOCATION_DESC)
data$LOCATION_DESC[data$LOCATION_DESC == ""] = NA
```

```{r summary, echo=FALSE}
summary(data)
```

This filtered data contains the location, district and exact time (if available) of the shooting incidents between 2006 and 2021, collected by the NYPD.

## Timing of incitents

```{r vis2, echo=FALSE}
barplot(table(data$OCCUR_DATETIME$hour))
```

We can clearly see that most incidents occurred during the late afternoon and evening hours, while the mornings are relatively safer, which is not that surprising.

```{r vis1, echo=FALSE}
barplot(table(data$OCCUR_DATETIME$year + 1900))
```

It looks like the amount of shooting incidents were on a decline, but right around when COVID started spread around the globe, the number of incidents almost doubled and the downward trend seems to be missing since then. This is a clear correlation between the pandemic and the amount shooting incidents.
I suspect that the amount of shooting incidents per year will stagnate while the pandemic is still ongoing and may start to decrease following that.

## Incident placements

```{r destricts, echo=FALSE}
pie(table(data$BORO))
```

Most people would say that Queens is the most dangerous, but turns out Brooklyn is where the most recorded shooting incidents happened.

```{r placemets, echo=FALSE}
sort(table(data$LOCATION_DESC), decreasing=TRUE)
```

It seams that most of the shootings occur in places where lots of people live close together like public houses and apartment buildings.

### Session information

```{r session_info, echo=FALSE}
s = sessionInfo()
print(s, locale=FALSE)
```